Assignment 1: Data and Descriptive

### 2023-11-09 MSB104_GR_1_Final_Assignment_research_article.qmd
update 09.11

---
title: "Regional Inequalities and economic growth; cross section and times series analysis"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
editor: visual
bibliography: [Regional_inequality_23.bib]
output-file: Assignment1.qmd
page-layout: full
number-sections: true
papersize: a4
---

>>>>

---
title: "Regional Inequalities and economic growth; cross section and times series analysis"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
editor: visual
bibliography: [Regional_inequality_23.bib, econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
csl: apa.csl
---

### 2023-11-09 assignment_1.qmd
update 09.11

---
title: "Assignment 1: Data and Descriptive"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations."
editor: visual
bibliography: [Econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
---

@@

library(tidyverse)
library(tinytex)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
```

@@

@@

# References

>>>>

---
title: "Assignment 1: Data and Descriptive"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations."
editor: visual
bibliography: [Econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
csl: apa.csl
---

@@

library(tidyverse)
library(tinytex)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
library(huxtable)
```

Må fikse kilder!!

@@

... p. 116

@@

# Assignment 2: Cross-sectional Estimates

## 2A: Growth and Inequity

### 1. Model Estimation

$\text{GDP per capita = } \beta_{0} + \beta_{1} \cdot GINI + \in$

# Making a dataframe that only contains observations from 2010, also filter the data so that gini is bigger than 0.

data2010 <- gini_nuts_2 %>% filter(year == 2010, gini_n2 > 0)
```

```{r}
#| echo: false
#| fig-cap: Descriptive statistics all countries for year 2010

# Finding the natural logarithm of GDP per capita
data2010 <- data2010 |> 
  mutate(log_GDP_per_capita = log(GDP_per_capita))

# lm(y ~ x) y = GDP per capita, x = Gini, estimation
lm_centered <- lm(gini_n2 ~ log_GDP_per_capita, data = data2010)

#Save coefficients
b0_earn <- lm_centered$coefficients[1]
b1_earn <- lm_centered$coefficients[2]

# Descriptive statistics
huxreg(All = lm_centered, statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
```

```{r}
#| echo: false
#| output: false
# må forklare denne!!
plot(lm_centered, which = 1:4)
```

β0​ = 0.0551 (intercept)

β1 = 0.0193 (slope coefficient)

```{r}
#| echo: false
#| fig-cap: Descriptive statistics for all countries seperately for the year 2010

# Finding estimates for each country seperately 
lmSK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "SK"))
lmDK <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "DK"))
lmHU <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "HU"))
lmPT <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "PT"))
lmFR <- lm(gini_n2 ~ log_GDP_per_capita, subset(data2010, id_nuts0 == "FR"))

# Descriptive statistics
huxreg(SK = lmSK, DK = lmDK, HU = lmHU, PT = lmPT, FR = lmFR, statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
```

### 2. Model Diagnostics

Forklaring på statistikk for alle land samlet:

Forklaring på statistikk for land hver for seg:

### 3. Visualization

# Visualization of the relationship between GDP per capita and gini
ggplot(data2010, aes(x = gini_n2, y = log_GDP_per_capita)) +
  geom_point(aes(color = gini_n2), size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship Between GDP per Capita and Gini",
    x = "Gini",
    y = "GDP per capita",
    caption = "Red line indicates the fitted regression line"
  ) +
  theme_minimal()

```

Forklaring på visualisering:

Forklar hvorfor noen observasjoner er langt unna de andre observasjonene.

## 2B: Exploring Other Determinants of Inequity

### 1. Data Acquistion

```{r}
#| echo: false
#| output: false
#| label: data acquistion

# Data for population density
demo_dens_data <- read_csv("Raw_data/demo_dens.gz")

# Data for education
educ_data <- read_csv("Raw_data/educ.gz")

# Data for rail network
rail_data <- read_csv("Raw_data/rail.gz")
```

# Cleaning up data frames
demo_dens_data <- demo_dens_data |> 
  select(nuts_2 = geo, pop_density = OBS_VALUE)

educ_data <- educ_data |> 
  select(nuts_2 = geo, students_percentage = OBS_VALUE)

rail_data <- rail_data |> 
  select(nuts_2 = geo, rail_km = OBS_VALUE)
```

```{r}
merged_data_1 <- demo_dens_data %>%
  inner_join(educ_data, by = 'nuts_2') %>%
  inner_join(rail_data, by = 'nuts_2')
```

#### Population Density

#### Education

#### Rail Network

### 2. Multiple Linear Regression Model

### 3. Model Interpretation
